---
title: "EP_Cereal"
author: "Project with three other members, Canberk Atak"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
getwd()
```

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(readxl)
library(data.table)
library(haven)
library(modelsummary)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(naniar)
library(fixest)
library(lfe)
library(sandwich)
library(lmtest)
library(knitr)
library(corrplot)

```

#Import data:

```{r}
datapath <- file.path(getwd(), "Cereal_Data.xls") 
cerealdata <- read_excel(datapath)
```

# TASK 1 

1\. **Data Description.** **Provide a brief, non-technical, overview of
the market using descriptive tools (tables, plots) and focusing on a lay
audience. Which are the salient features of the market, the main brands,
segments, and products?**

Missing observations:

```{r}
sapply(cerealdata, function(x) sum(is.na(x)))
```

```{r}
library(naniar)
vis_miss(cerealdata, warn_large_data = FALSE)
```

In total 1% of observations are missing values.

## Remove rows with missing values

```{r}
cerealdata <- na.omit(cerealdata)
```

```{r}
datasummary( All(cerealdata) ~ mean + sd + min + median + max, fmt = 4, data = cerealdata, output = "markdown")
```

##discount&coupons

```{r}
# create a variable "coupon"
cerealdata <- cerealdata%>%
  mutate(coupon = `Avg Shelf Price` - `Avg Trans Price`)
```

```{r}

# Group, breakfast cereals under the different market segment 
avg_prices_by_segment <- cerealdata %>%
  group_by(Sgmnt) %>%
  summarise(AvgShelfPrice = mean(`Avg Shelf Price`, na.rm = TRUE))

print(avg_prices_by_segment)

# Different prices for consumer heterogeneity

plot_1 <- ggplot(cerealdata, aes(x = Sgmnt, y = `Avg Shelf Price`, fill = Sgmnt)) +
  geom_boxplot() +
  labs(x = "Segment",
       y = "Average Shelf Price") +
  theme_minimal() +
  scale_fill_brewer(palette = "Pastel1") + 
  theme(plot.title = element_text(face = "italic"))

ggsave("Comparison of Average Shelf Prices by Segment.jpeg", plot = plot_1, width = 10, height = 6, dpi = 600)

print(plot_1)

```

```{r}
# 
# # Top 10 cereals in the market based on their Market share
# 
# plot_2 <- cerealdata %>%
#   top_n(10, `Mkt Share`) %>%
#   ggplot(aes(x = reorder(Name, `Mkt Share`), y = `Mkt Share`)) +
#   geom_bar(stat = "identity", fill = "darkblue") +
#   coord_flip() + # This makes it horizontal for better readability
#   labs(title = "Top 10 Cereals by Market Share", x = "", y = "Market Share (%)") +
#   theme_minimal()
# 
# print(plot_2)
# 
# ggsave(filename = "Top 10 Cereals by Market Share.jpeg" ,plot_2 )

```

```{r}
# Seeing the different price and different nutritional values
# Identifying the row with the maximum value for each variable

max_fat <- cerealdata[which.max(cerealdata$Fat),]
max_cals <- cerealdata[which.max(cerealdata$Cals),]
max_sugar <- cerealdata[which.max(cerealdata$Sugar),]
max_price <- cerealdata[which.max(cerealdata$`Avg Shelf Price`),]

# Combining these into a single data frame for annotation
max_values <- bind_rows(max_fat, max_cals, max_sugar, max_price)

plot_3 <- ggplot(cerealdata, aes(x = Sugar, y = `Avg Shelf Price`, size = Cals, color = Fat)) +
  geom_point(alpha = 0.7) +
  geom_label(data = max_values, 
             aes(label = paste("F:", round(Fat,1), "C:", round(Cals,0), "S:", round(Sugar,0), "P:", round(`Avg Shelf Price`,2))), 
             hjust = 0.9, vjust = 1.5, 
             size = 3,
             label.size = NA,
             lineheight = 0.9, 
             fontface = "italic", 
             color = "black") +
  scale_color_gradient(low = "yellow", high = "red") +
  labs(x = "Sugar Content (S)", 
       y = "Avg Shelf Price (P)", 
       color = "Fat Content", 
       size = "Calorie Count") +
  theme_minimal() +
  guides(color = guide_colourbar(title = "Fat (F)"), size = guide_legend(title = "Calories (C)")) +
  theme(legend.position = "right")


ggsave(filename = "Average Shelf Price vs. Nutritional Values.jpeg", plot = plot_3, dpi = 900)
print(plot_3)

```

```{r}
# ggplot(cerealdata, aes(x = `Avg Ad Expn`, y = coupon)) +
#   geom_point(aes(color = Sgmnt), alpha = 0.7) + 
#   geom_smooth(method = "lm", se = FALSE, color = "gray") +
#   labs(title = "Relationship Between Advertising Expenses and Discounts",
#        x = "Average Advertising Expenditure",
#        y = "Discount Amount") +
#   theme_minimal() +
#   scale_color_brewer(palette = "Pastel1")

```

```{r}
# Different Discounts for different Segment

plot_4 <- ggplot(cerealdata, aes(x = Sgmnt, y = coupon, fill = Sgmnt)) +
  geom_point() +
  labs(x = "Segment",
       y = "Discount Amount") +
  theme_minimal() +
  scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.5) +
  theme(plot.title = element_text(face = "italic"))


ggsave(filename = "Discount Distribution by Segment.jpeg",plot = plot_4, dpi = 600)

print(plot_4)

```

```{r}
# 

plot_5 <- cerealdata %>%
  group_by(Sgmnt) %>%
  summarise(AvgCalories = mean(Cals, na.rm = TRUE),
            AvgFat = mean(Fat, na.rm = TRUE),
            AvgSugar = mean(Sugar, na.rm = TRUE)) %>%
  gather(key = "Nutrient", value = "Average", -Sgmnt) %>%
  ggplot(aes(x = Sgmnt, y = Average, fill = Nutrient)) +
  geom_bar(stat = "identity") +
  labs(x = "Segment", y = "Average Value") +
  theme_classic() +
  scale_fill_brewer(palette = "Pastel1")


ggsave(filename = "Average Nutritional Values by Segment.jpeg",plot = plot_5, dpi = 600)
print(plot_5)
```

```{r}
unique_product <- unique(cerealdata$Name)
unique_product
```

##brand

```{r}
cerealdata <- cerealdata %>%
  mutate(Brand = sub(" .*", "", Name))
```

```{r}
unique_brand <- unique(cerealdata$Brand)
unique_brand
```

## Plots of market share grouped by brand

```{r}

# Calculating total market share for each brand
df_brand <- cerealdata %>%
  group_by(Brand) %>%
  summarize(total_market_share = sum(`Mkt Share`)) %>%
  arrange(desc(total_market_share))

# Plot
plot_6 <- ggplot(df_brand, aes(x = reorder(Brand, total_market_share), y = total_market_share, fill = Brand)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = total_market_share), position = position_stack(vjust = 0.5), size = 3.5, color = "black") +
  labs(x = "Brand",
       y = "Total Market Share (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

ggsave("Market Share by Brands.jpeg", plot_6)
print(plot_6)

```

```{r}

# Counting the number of products for each brand
brand_counts <- cerealdata %>%
  group_by(Brand) %>%
  summarise(Products = n())

# Creating a bar plot showing the number of products for each brand
plot_7 <- ggplot(brand_counts, aes(x = reorder(Brand, Products), y = Products, fill = Brand)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Products), position = position_stack(vjust = 0.5), size = 3.5, color = "black") +
  coord_flip() +
  labs(x = "Brand",
       y = "Number of Products") +
  theme_minimal() +
  theme(legend.position = "none")

ggsave("Number of Products by Brand.jpeg", plot_7)

print(plot_7)


```

```{r}
unique_product <- unique(cerealdata$Name)
unique_product
```

## Plots of market share grouped by brand

```{r}
library(ggplot2)

# Calculate total market share for each brand
df_brand <- cerealdata %>%
  group_by(Brand) %>%
  summarize(total_market_share = sum(`Mkt Share`)) %>%
  arrange(desc(total_market_share))

# Plot
ggplot(df_brand, aes(x = reorder(Brand, total_market_share), y = total_market_share, fill = Brand)) +
  geom_bar(stat = "identity") +
  labs(title = "Market Share by Brand",
       x = "Brand",
       y = "Total Market Share (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

```

# TASK 2

2.  **The basics and OLS estimation.** **Load the data, define the share
    of the outside good, create a variable to store the brand
    information and create the dependent variable to be used in the
    logit regressions (mean utility). Then estimate the following six
    specifications using OLS and including product characteristics
    Price, Cals, Fat, and Sugar:**

Load data:

```{r}
datapath <- file.path(getwd(), "Cereal_Data.xls") 
cerealdata <- read_excel(datapath)
cerealdata <- na.omit(cerealdata)
names(cerealdata)
```

*share of the outside good*

```{r}
outside_share=100-round(sum(cerealdata$`Mkt Share`))
```

*brand information*

```{r}
cerealdata <- cerealdata %>%
  mutate(
    Brand = word(Name, 1, sep = " "),
    Product = word(Name, 2, sep = " "))
cerealdata$fbrand <- as.factor(cerealdata$Brand)
cerealdata$fsegment <- as.factor(cerealdata$Sgmnt)
```

```{r}
names(cerealdata)
```

```{r}
cerealdata <- cerealdata %>%
  mutate(dummy = model.matrix(~ fbrand - 1, data = .))

```

*mean utility as dependent variable*

```{r}
cerealdata <- cerealdata %>%
  mutate(mean_utility = log(`Mkt Share`) - log(outside_share))
      
```

**Then estimate the following six specifications using OLS and including
product characteristics Price, Cals, Fat, and Sugar:**

```{r}
names(cerealdata)

```

*(1) Without brand fixed-effects, standard errors assuming IID errors;*

```{r}
# Model specification without brand fixed-effects and with IID standard errors
ols_1 <- feols(mean_utility ~Cals + Fat + Sugar+`Avg Shelf Price`, data = cerealdata, se = 'iid')
summary(ols_1)



```

*(2) Without brand fixed-effects, heteroskedasticity-robust standard
errors;*

```{r}
ols_2 <- feols(mean_utility ~ Cals + Fat + Sugar+`Avg Shelf Price`, data = cerealdata, vcov = "HC1")
summary(ols_2)
```

*(3) Without brand fixed-effects, standard errors clustered at the brand
level*

```{r}

ols_3 <- feols(mean_utility ~ Cals + Fat + Sugar+`Avg Shelf Price` , data = cerealdata, se = 'cluster', cluster = "Brand")
summary(ols_3)


```

*By adding a brand fixed-effect I control for unobserved quality,* *(4)
With brand fixed-effects, standard errors assuming IID errors;*

```{r}

ols_4 <- feols(mean_utility ~ Cals + Fat + Sugar+`Avg Shelf Price`| fbrand, data = cerealdata,
                     se = 'iid')
summary(ols_4)
```

*(5) With brand fixed-effects, heteroskedasticity-robust standard
errors*

```{r}
ols_5 <- feols(mean_utility ~ Cals + Fat + Sugar+`Avg Shelf Price`| fbrand, data = cerealdata,
                     vcov = 'HC1')
summary(ols_5)
```

*(6) With brand fixed-effects, standard errors clustered at the brand
level*

```{r}
ols_6 <- feols(mean_utility ~ Cals + Fat + Sugar+`Avg Shelf Price` | fbrand,  data = cerealdata)
summary(ols_6)
```

#table with results#

```{r}
coef <- list('Intercept'="Intercept",
              'Cals' = 'Caloric content of the product, per portion',
             'Fat' = 'Fat content of the product, per portion',
             'Sugar' = 'Sugar content of the product, per portion',
             'Avg Shelf Price' = 'Average shelf price')

gof <- data.frame(raw = c("nobs", "adj.r.squared"),
                  clean = c("Observations", "R2 adj."),
                  fmt = c(0,3))

fe <- data.frame(FE = c("Brand FE","errors"),
                 mod1 = c("No","idd"),
                 mod2 = c("No","HC1"),
                 mod3 = c("No","clustered"),
                 mod4 = c("Yes","idd"),
                 mod5 = c("Yes","HC1"),
                 mod6 = c("Yes","clustered"))
attr(fe, "position") <- c(11,12)

modelsummary(list(ols_1, ols_2, ols_3, ols_4, ols_5,ols_6),
              stars = c("*" = .1, "**" = .05, "***" = .01),
              title = "Table 1: Analysis of cereal market using OLS with and without brand fixed-effects",gof_map = gof, add_rows = fe, output = "markdown")

```

```{r}
cerealdata <- cerealdata %>%
  mutate(discount_coupon = `Avg Shelf Price` - `Avg Trans Price`)

```

# TASK 3

3.  **Instruments and IV estimation. Construct instruments to account
    for price endogeneity and use them to perform IV estimation. Report
    estimates on a table with specifications differing according to the
    instruments used. Which instruments have you constructed, and why?
    Can one claim that the more the merrier when it comes to instruments
    in this case? Note. Use the OLS estimates without and with brand
    fixed-effects with your preferred standard errors. Instrument each
    of them with the instruments you constructed and report a table with
    6-8 specifications in total**.

```{r}
library(ggplot2)
library(corrplot)
library(openxlsx)
cerealdata <- cerealdata %>%
  mutate(fsgmnt = as.factor(Sgmnt))
cor_matrix <- cor(subset(cerealdata, select = c("Avg Shelf Price", "Avg Trans Price" ,"Avg Ad Expn")))
print(cor_matrix)


correlation_matrix <- cor(cerealdata[, c("Avg Shelf Price","Avg Trans Price","Avg Ad Expn")])

png("heatmap_base.png", width = 800, height = 1200, res = 300)
heatmap(correlation_matrix)
dev.off()


png("heatmap_corrplot.png", width = 800, height =1080, res =150)
corrplot(correlation_matrix, method = "color",
         type = "upper", 
         order = "hclust", 
         addCoef.col = "black", 
         col = colorRampPalette(c("lightblue", "darkblue"))(200), 
         title = "Correlation Matrix", 
         tl.col = "black", tl.srt = 30) 
dev.off()





```

```{r}
#  OLS with fixed effects
summary(ols1_1 <- feols(mean_utility ~ Cals + Fat + Sugar + `Avg Shelf Price`, data = cerealdata, se = 'iid'))

#IV with average advertising expenditure * brand
summary(iv1a_2 <- feols(mean_utility ~ Cals + Fat + Sugar | `Avg Shelf Price` ~ `Avg Ad Expn`:fbrand, data = cerealdata, se = 'iid'))

# IV with average average transaction price * brand
summary(iv1a_3 <- feols(mean_utility ~ Cals + Fat + Sugar  | `Avg Shelf Price` ~ `Avg Trans Price`:fbrand, data = cerealdata, se = 'iid'))

#  IV with average average transaction price * brand + advertising expenditure * brand
summary(iv1a_4 <- feols(mean_utility ~ Cals + Fat + Sugar | `Avg Shelf Price` ~ `Avg Trans Price`:fbrand + `Avg Ad Expn`:fbrand, data = cerealdata, se = 'iid'))

#  benchmark: OLS with fixed effects
summary(ols_4_1 <- feols(mean_utility ~ Cals + Fat + Sugar + `Avg Shelf Price` | fbrand, data = cerealdata,
                     se = 'iid'))

#IV with average advertising expenditure * brand
summary(iv2a_5 <- feols(mean_utility ~ Cals + Fat + Sugar | fbrand | `Avg Shelf Price` ~ `Avg Ad Expn`:fbrand,
                        data = cerealdata, se = 'iid'))

#  IV with  average transaction price * brand
summary(iv2a_6 <- feols(mean_utility ~ Cals + Fat + Sugar | fbrand | `Avg Shelf Price` ~ `Avg Trans Price`:fbrand, 
                        data = cerealdata, se = 'iid'))

#  IV with both  average transaction price * brand and advertising expenditure * brand
summary(iv2a_7 <- feols(mean_utility ~ Cals + Fat + Sugar | fbrand |
                               `Avg Shelf Price` ~ `Avg Ad Expn`:fbrand + `Avg Trans Price`:fbrand,
                             data = cerealdata, se = 'iid'))

# Define coefficients
coef <- list('Intercept'="Intercept",
             'Cals' = 'Caloric content of the product, per portion',
             'Fat' = 'Fat content of the product, per portion',
             'Sugar' = 'Sugar content of the product, per portion',
             'Avg Shelf Price' = 'Average shelf price')

# Define goodness-of-fit dataframe
gof <-  data.frame(raw = c("nobs", "adj.r.squared"),
                   clean = c("Observations", "R2 adj."),
                   fmt = c(0,3))

# Define fixed effects dataframe
fe <- data.frame(FE = c("Brand FE","Instruments"),
                 mod1 = c("No","None"),
                 mod2 = c("No", "Av. Adv Exp x Brand"),
                 mod3 = c("No","Avg Trans Price x Brand"),
                 mod4 = c("No","Avg Trans Price + Avg Trans Price x Brand"),
                 mod5 = c("Yes","None"),
                 mod6 = c("Yes","Av. Adv Exp x Brand"),
                 mod7 = c("Yes","Sugar x Brand"),
                 mod8 = c("Yes", "Avg Trans Price  + Avg Trans Price x Brand"))
                  
attr(fe, "position") <- c(14,15)

# Generate model summary in markdown format
modelsummary(list(ols1_1, iv1a_2, iv1a_3, iv1a_4, ols_4_1, iv2a_5, iv2a_6,iv2a_7),
             stars = c("*" = .1, "**" = .05, "***" = .01),
             output = "markdown")

```

## Testing the Instrumental variables:

```{r}
fitstat(iv1a_2, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv1a_3, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv1a_4, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv2a_5, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv2a_6, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv2a_7, ~ ivwald1 + ivf1 + kpr + sargan)
```

# TASK 4

**More instruments and IV estimation. In case you didn’t get a
satisfactory specification, can you combine instruments to get better
results? If so, how? Report a table with your estimates.**

```{r}
#price of other products within the same brand
cerealdata <- cerealdata %>%
  group_by(fbrand) %>%
  mutate(brand_av_prices = sum(`Avg Shelf Price`))

# Calculate iva variable
cerealdata$iva <- with(cerealdata, {
  ifelse(Brand == "GM", (1/18) * (brand_av_prices-`Avg Shelf Price`  ),
  ifelse(Brand == "KG", (1/17) * (brand_av_prices-`Avg Shelf Price` ),
  ifelse(Brand == "NB", (1/3) * (brand_av_prices-`Avg Shelf Price` ),
  ifelse(Brand == "PT", (1/6) * (brand_av_prices-`Avg Shelf Price` ),
  ifelse(Brand == "QK", (1/4) * (brand_av_prices-`Avg Shelf Price` ),
  ifelse(Brand == "RL", (1/2) * (brand_av_prices-`Avg Shelf Price` ),
         (brand_av_prices-`Avg Shelf Price` )))))))})

# Compute correlation

cor(subset(cerealdata, select = c("Avg Shelf Price", "iva")))
```

```{r}
# discount coupon
cerealdata <- cerealdata %>%
  mutate(ivb = discount_coupon)

cor(subset(cerealdata, select = c("Avg Shelf Price", "ivb")))


```

```{r}
summary(cerealdata$fsegment)

```

```{r}
# price by segment
cerealdata <- cerealdata %>%
  group_by(fsegment) %>%
  mutate(segment_av_prices = sum(`Avg Shelf Price`))

cerealdata$ivc <- with(cerealdata, {
  ifelse(fsegment == "Adult", (1/25) * (segment_av_prices - `Avg Shelf Price`  ),
  ifelse(fsegment == "Fam", (1/13) * (segment_av_prices - `Avg Shelf Price` ),
  ifelse(fsegment == "Kids", (1/12) * (segment_av_prices - `Avg Shelf Price` ),
         (segment_av_prices - `Avg Shelf Price` ))))})

# Compute correlation

cor(subset(cerealdata, select = c("Avg Shelf Price", "ivc")))
```

```{r}
cor(subset(cerealdata, select = c("Avg Shelf Price", "iva", "ivb", "ivc")))
```

```{r}
# Compute correlation matrix
corr <- cor(subset(cerealdata, select = c("Avg Shelf Price", "iva", "ivb", "ivc")))

# Plot heatmap with variable names
heatmap(corr, labRow = c("Avg Shelf Price", "iva", "ivb", "ivc"), 
        labCol = c("ASPrice", "iva", "ivb", "ivc"))

```

```{r}
#model
summary(ols2_1 <- feols(mean_utility ~ Cals + Fat + Sugar +`Avg Shelf Price`, data = cerealdata,
                     se = 'iid'))
summary(iv3c_1 <- feols(mean_utility ~ 1 |  `Avg Shelf Price`~ iva, data = cerealdata,
                     se = 'iid'),stage = 1:2)
summary(iv3c_2 <- feols(mean_utility ~ 1 |  `Avg Shelf Price`~ ivb, data = cerealdata,
                     se = 'iid'),stage = 1:2)
summary(iv3c_3 <- feols(mean_utility ~ 1 |  `Avg Shelf Price`~ ivc, data = cerealdata,
                     se = 'iid'),stage = 1:2)

summary(ols2_2 <- feols(mean_utility ~ Cals + Fat + Sugar +`Avg Shelf Price`|fbrand, data = cerealdata,
                     se = 'iid'))

summary(iv3c_4 <- feols(mean_utility ~ fbrand |  `Avg Shelf Price`~ iva, data = cerealdata,
                     se = 'iid'),stage = 1:2)
summary(iv3c_5 <- feols(mean_utility ~ fbrand |  `Avg Shelf Price`~ ivb, data = cerealdata,
                     se = 'iid'),stage = 1:2)
summary(iv3c_6 <- feols(mean_utility ~ fbrand |  `Avg Shelf Price`~ ivc, data = cerealdata,
                     se = 'iid'),stage = 1:2)



gof <- data.frame(raw = c("nobs", "adj.r.squared"),
                  clean = c("Observations", "R2 adj."),
                  fmt = c(0, 3))
fe <- data.frame(FE = c("Brand FE", "Instruments"),
                 mod1 = c("No", "None"),
                 mod2 = c("No", "IVA"),
                 mod3 = c("No", "IVB"),
                 mod4 = c("No", "IVC"),
                mod5 = c("Yes", "None"),
                mod6 = c("Yes", "IVA"),
                mod7 = c("Yes", "IVB"),
                mod8 = c("Yes", "IVC"))

attr(fe, "position") <- c(23, 24)

modelsummary(list(ols2_1, iv3c_1,iv3c_2,iv3c_3, ols2_2,iv3c_4,iv3c_5,iv3c_6),
              stars = c("*" = .1, "**" = .05, "***" = .01),
              gof_map = gof, add_rows = fe, output = "markdown")

```

```{r}
fitstat(iv3c_1, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv3c_2, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv3c_3, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv3c_4, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv3c_5, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv3c_6, ~ ivwald1 + ivf1 + kpr + sargan)
```

#interaction

```{r}
summary(ols2_1 <- feols(mean_utility ~ Cals + Fat + Sugar +`Avg Shelf Price`, 
                        data = cerealdata, se = 'iid'))

summary(iv4c_1 <- feols(mean_utility ~ 1 |  `Avg Shelf Price`~ iva +  ivb + ivc, 
                        data = cerealdata, se = 'iid'),stage = 1:2)

summary(ols2_2 <- feols(mean_utility ~ Cals + Fat + Sugar +`Avg Shelf Price`|fbrand, 
                        data = cerealdata,se = 'iid'))

summary(iv4c_2 <- feols(mean_utility ~ fbrand |  `Avg Shelf Price`~ iva + ivb +ivc, 
                        data = cerealdata, se = 'iid'),stage = 1:2)

gof <- data.frame(raw = c("nobs", "adj.r.squared"),
                  clean = c("Observations", "R2 adj."),
                  fmt = c(0, 3))
fe <- data.frame(FE = c("Brand FE", "Instruments"),
                 mod1 = c("No", "None"),
                 mod2 = c("No", "IVA+IVB+IVC"),
                 mod3 = c("Yes", "None"),
                 mod4 = c("Yes", "IVA+IVB+IVC"))

attr(fe, "position") <- c(23, 24)

modelsummary(list(ols2_1, iv4c_1, ols2_2, iv4c_2),
              stars = c("*" = .1, "**" = .05, "***" = .01),
              gof_map = gof, add_rows = fe, output = "markdown")

```

```{r}
fitstat(iv4c_1, ~ ivwald1 + ivf1 + kpr + sargan)
```

```{r}
fitstat(iv4c_2, ~ ivwald1 + ivf1 + kpr + sargan)
```

# TASK 5

**Thinking out of the matrix. In case you didn’t get a satisfactory
specification up to now, can you suggest any improvements? Hint: One
improvement comes from a new variable. Another improvement comes from
tackling the IIA property of logit models. As a guidance, When
reporting, please include again the following specifications previously
reported: (i) your preferred OLS logit specifications with and without
brand FEs; (ii) your preferred IV logit specification with and without
brand FEs which include all instruments.**

The "IIA property" refers to the Independence of Irrelevant
Alternatives, which is a key assumption in multinomial logit models.
This assumption states that the odds of choosing one alternative over
another are unaffected by the presence or absence of other alternatives
in the choice set.

However, the IIA property is often criticized because it can lead to
unrealistic modeling assumptions. When the IIA property is violated, it
implies that the relative attractiveness of alternatives is influenced
by the inclusion or exclusion of irrelevant alternatives in the choice
set. This can result in biased model estimates and misleading
predictions.

There are several techniques to address or mitigate violations of the
IIA property:

1.  **Nested Logit Models**: These models relax the assumption of
    independence by allowing for correlation within subsets of
    alternatives. Alternatives are grouped into nested subsets based on
    similarities or shared characteristics. This allows for more
    flexibility in capturing heterogeneity in preferences.

```{r}
# Create dummy variable for sugar
cerealdata$sugar_dummy <- ifelse(cerealdata$Sugar >= 1, 1, 0)

#withouth brand fix effect baseline model
ols_1o <- feols(mean_utility ~ Cals + Fat + Sugar + `Avg Shelf Price`, data = cerealdata,
                se = 'iid')
summary(ols_1o)

#new variable
ols_1n <- feols(mean_utility ~ Cals + Fat + Sugar + `Avg Shelf Price` + sugar_dummy, data = cerealdata,
                se = 'iid')
summary(ols_1n)


#with brand fix effect baseline model
ols_4o <- feols(mean_utility ~ Cals + Fat + Sugar + `Avg Shelf Price` | fbrand, data = cerealdata,
                se = 'iid')
summary(ols_4o)

#new variable
ols_4n <- feols(mean_utility ~ Cals + Fat + Sugar + `Avg Shelf Price` + sugar_dummy | fbrand, data = cerealdata,
                se = 'iid')
summary(ols_4n)



# without fe  iv1a_2 Average advert Expenditure*brand 

summary(iv1a_2o <- feols(mean_utility ~ Cals + Fat + Sugar | `Avg Shelf Price` ~ `Avg Ad Expn`:fbrand, data = cerealdata, se = 'iid'))

#without fe  iv1a_2 Average advert Expenditure*brand 
summary(iv1a_2n <- feols(mean_utility ~ Cals + Fat + Sugar + sugar_dummy | `Avg Shelf Price` ~ `Avg Ad Expn`:fbrand, data = cerealdata, se = 'iid'))



# with fe  iv1a_6 Average transaction price*brand 

summary(iv2a_6o <- feols(mean_utility ~ Cals + Fat + Sugar | fbrand | `Avg Shelf Price` ~ `Avg Trans Price`:fbrand, 
                        data = cerealdata, se = 'iid'))


# with fe  iv1a_6 Average transaction price*brand 

summary(iv2a_6n <- feols(mean_utility ~ Cals + Fat + Sugar +sugar_dummy| fbrand | `Avg Shelf Price` ~ `Avg Trans Price`:fbrand, 
                        data = cerealdata, se = 'iid'))


gof <- data.frame(raw = c("nobs", "adj.r.squared"),
                  clean = c("Observations", "R2 adj."),
                  fmt = c(0, 3))
fe <- data.frame(FE = c("Brand FE", "Instruments","New variable"),
                 mod1 = c("No", "None","No"),
                 mod2 = c("No", "None","Yes"),
                 mod3 = c("Yes", "None","No"),
                 mod4 = c("Yes", "None","Yes"),
                 mod5 = c("No", "Average advert Expenditure*brand","No"),
                 mod6 = c("No", "Average advert Expenditure*brand","Yes"),
                 mod7 = c("Yes", "Average Avg Trans Price*brand","No"),
                 mod8 = c("Yes", "Average Avg Trans Price*brand","Yes"))
                 

attr(fe, "position") <- c(15, 16)

modelsummary(list(ols_1o, ols_1n, ols_4o, ols_4n, iv1a_2o, iv1a_2n, iv2a_6o, iv2a_6n),
              stars = c("*" = .1, "**" = .05, "***" = .01),
              gof_map = gof, add_rows = fe, output = "markdown")

```

###QUESTION6

```{r}
summary(iv5 <- feols(mean_utility ~ fbrand | `Avg Shelf Price` ~ iva, data = cerealdata, se = 'iid'), stage = 1:2)

elas <- merge(aggregate(cerealdata$mean_utility, list(cerealdata$fbrand), FUN = mean),
              aggregate(cerealdata$`Avg Shelf Price`, list(cerealdata$fbrand), FUN = mean), by = "Group.1")
names(elas) <- c("fbrand", "mean_utility_", "Avg_Shelf_Price_")
elas$op_el <- coef(iv5)["fit_`Avg Shelf Price`"] * (elas$Avg_Shelf_Price_ / elas$mean_utility_)
datasummary(Heading("") * mean * fbrand ~ Heading("Own-price elasticity") * op_el,
             elas, fmt = 4, output = "markdown")

```

η = −αpj (1 − sj ), if j = k -\> so own price elasticities η = αpk sk ,
else (meaning j not equal to k) -\> cross price elasticities

```{r}
summary(iv2a_6)
```

```{r}
coefficients <- coef(iv2a_6)
print(coefficients)
```

η = −αpj (1 − sj ), if j = k -\> so own price elasticities η = αpk sk ,
else (meaning j not equal to k) -\> cross price elasticities

```{r}
cerealdata$`Mkt Share`
```

```{r}
# Extract coefficients
coefficients <- coef(iv2a_6)

# Calculate own-price elasticity (j = k)
own_price_elasticity <- -coefficients["fit_`Avg Shelf Price`"] * cerealdata$`Avg Shelf Price` * (1 - (cerealdata$`Mkt Share`/100))

# Calculate cross-price elasticity (j ≠ k)
cross_price_elasticity <- coefficients ["fit_`Avg Shelf Price`"]* cerealdata$`Avg Shelf Price` * (cerealdata$`Mkt Share`/100)


print("Own-price elasticity:")
print(own_price_elasticity)
print("Cross-price elasticity:")
print(cross_price_elasticity)

```

#Plot own price elasticities against prices

```{r}
library(ggplot2)

# Create a data frame with the own price elasticities and prices
elasticity_df <- data.frame(
  Avg_Shelf_Price = cerealdata$`Avg Shelf Price`,
  Own_Price_Elasticity = own_price_elasticity
)

# Plot own price elasticities against prices using ggplot2
plot_elas_1 <- ggplot(elasticity_df, aes(x = Avg_Shelf_Price, y = Own_Price_Elasticity)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Average Shelf Price", y = "Own-price Elasticity") +
  theme_minimal()

ggsave("Own price elasticities against price.jpeg", plot = plot_elas_1, dpi = 900 )

print(plot_elas_1)


```

#Plot markups against prices

```{r}
library(ggplot2)
library(dplyr)

# Calculate markup as a percentage
cerealdata <- cerealdata %>%
  mutate(markup_percentage = discount_coupon * 100)

# Plot markups against prices using ggplot2
plot_elas_2 <- ggplot(cerealdata, aes(x = `Avg Shelf Price`, y = markup_percentage)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Average Shelf Price", y = "Markup (%)") +
  theme_minimal()

ggsave("Markups against prices.jpeg", plot = plot_elas_2, dpi = 900)

print(plot_elas_2)

```

#matrix 50X 50

```{r}
# Extract coefficients
coefficients <- coef(iv2a_6)

# Get the number of products
num_products <- nrow(cerealdata)

# Initialize an empty matrix to store the elasticities
elasticity_matrix <- matrix(0, nrow = num_products, ncol = num_products)

# Calculate own-price elasticity (j = k) and cross-price elasticity (j ≠ k)
for (j in 1:num_products) {
  for (k in 1:num_products) {
    if (j == k) {
      # Calculate own-price elasticity
      elasticity_matrix[j, k] <- -coefficients["fit_`Avg Shelf Price`"] * cerealdata$`Avg Shelf Price`[j] * (1 - (cerealdata$`Mkt Share`[j]/100))
    } else {
      # Calculate cross-price elasticity
      elasticity_matrix[j, k] <- coefficients["fit_`Avg Shelf Price`"] * cerealdata$`Avg Shelf Price`[j] * (cerealdata$`Mkt Share`[k]/100)
    }
  }
}

write.csv(elasticity_matrix, file = "elasticity_matrix.csv", row.names = FALSE)

# Print the elasticity matrix
print(elasticity_matrix)



```

```{r}
# Load the xtable package
library(xtable)

# Extract coefficients
coefficients <- coef(iv2a_6)

# Get the number of products
num_products <- nrow(cerealdata)

# Initialize an empty matrix to store the elasticities
elasticity_matrix <- matrix(0, nrow = num_products, ncol = num_products)

# Calculate own-price elasticity (j = k) and cross-price elasticity (j ≠ k)
for (j in 1:num_products) {
  for (k in 1:num_products) {
    if (j == k) {
      # Calculate own-price elasticity
      elasticity_matrix[j, k] <- -coefficients["fit_`Avg Shelf Price`"] * cerealdata$`Avg Shelf Price`[j] * (1 - (cerealdata$`Mkt Share`[j]/100))
    } else {
      # Calculate cross-price elasticity
      elasticity_matrix[j, k] <- coefficients["fit_`Avg Shelf Price`"] * cerealdata$`Avg Shelf Price`[j] * (cerealdata$`Mkt Share`[k]/100)
    }
  }
}

# Create a LaTeX table
latex_table <- xtable(elasticity_matrix, caption = "Elasticity Matrix")

# Write the LaTeX table to a file
print(latex_table, file = "elasticity_matrix.tex")

```

```{r}
# Extract coefficients
coefficients <- coef(iv2a_6)

# Get the number of products
num_products <- nrow(cerealdata)

# Initialize an empty matrix to store the elasticities
elasticity_matrix <- matrix(0, nrow = num_products, ncol = num_products)

# Calculate own-price elasticity (j = k) and cross-price elasticity (j ≠ k)
for (j in 1:num_products) {
  for (k in 1:num_products) {
    if (j == k) {
      # Calculate own-price elasticity
      elasticity_matrix[j, k] <- -coefficients["fit_`Avg Shelf Price`"] * cerealdata$`Avg Shelf Price`[j] * (1 - (cerealdata$`Mkt Share`[j]/100))
    } else {
      # Calculate cross-price elasticity
      elasticity_matrix[j, k] <- coefficients["fit_`Avg Shelf Price`"] * cerealdata$`Avg Shelf Price`[j] * (cerealdata$`Mkt Share`[k]/100)
    }
  }
}

# Convert the matrix to a data frame for xtable
df_elasticity <- as.data.frame(elasticity_matrix)

# Creating LaTeX code for the table
latex_table <- xtable(df_elasticity, caption = "Elasticity Matrix", label = "tab:elasticity_matrix")

# Printing the LaTeX code for the table
print(latex_table, type = "latex", include.rownames = FALSE, 
      sanitize.text.function = function(x){x},
      floating = FALSE, 
      hline.after = c(-1, 0, nrow(df_elasticity)), 
      booktabs = TRUE)

# Saving the LaTeX code to a file
writeLines(capture.output(print(latex_table, type = "latex", include.rownames = FALSE, 
                                sanitize.text.function = function(x){x},
                                floating = FALSE,
                                hline.after = c(-1, 0, nrow(df_elasticity)),
                                booktabs = TRUE)),
           "elasticity_matrix.tex")



```
